The Courtroom Challenge—A Conjecture on the Unavoidable Cost of LLM Verification

### 2025-07-16 main.tex
Add files via upload
Generalize from text‐only to modality‐agnostic “data streams” + entropy cleanup

* **Global scope**
  * Replace “token/query stream” → “open‑world data stream” (variables Q→S).
  * Swap uniform‑prior constant log M for entropy placeholder H_* throughout.

* **Theory section**
  * Lemma ErrLower: now stated for arbitrary prior; boxed bound uses H_*.
  * Added un‑numbered *Remark* explaining H_* convention (Shannon vs min‑entropy);
    removed obsolete footnote.
  * Proof + discussion: log M → H_*; γ defined with H_*.

* **Challenge & narrative**
  * tcolorbox *Falsify‑Us*:

% ---------- math & symbols ----------
\usepackage{amsmath,amssymb,amsfonts}

@@

% =========================================================
\begin{document}
\maketitle
\begin{abstract}
  Large language models hallucinate; ad-hoc patches conceal the true cost of verification. 
  We therefore pose a falsifiable ultimatum: either automate validation, prove it unnecessary, or pay the explicit human-in-the-loop cost.  
  Conjecture 1 (the Courtroom Challenge). 
  For any unbounded open-world text stream, a system that lacks both (i) a persistent precedent store and (ii) a friction-priced validation loop will accumulate unbounded prequential regret.
  We formalise the state machine, cost ledger, and impossibility triangle that make this conjecture testable, and invite the community to break it.
  A 100-query toy stream illustrates how the ledger is logged\footnote{Baseline script will appear at \url{https://github.com/johnrepsys/courtroom-model-lean} with v2.}; no attempt is made here to optimise the validator.
  In addition, we present the Courtroom Model, a wrapper architecture that subjects every LLM claim to a perpetual cycle of precedent search, automated semantic checks, and selective escalation to an external judge—human or automated.
  Each new output is validated against a persistent vector memory, while conflicting evidence triggers the re-examination of prior “case law.”
  The challenge stands: thrive in an open world without paying the validation bill—or show why the bill is inescapable.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Large Language Models (LLMs) have demonstrated impressive capabilities in generating context-rich outputs by associating patterns from large corpora. However, this power comes with a critical limitation: LLMs often produce plausible-sounding but factually incorrect conclusions, a phenomenon known as hallucination \citep{marcus2020nextdecade,shinn2023reflexion}. Without structured verification, these errors risk being perpetuated and even amplified over time, undermining trust in LLM-generated knowledge. Addressing this issue requires a system that not only generates associations but rigorously validates and refines them.

To address these shortcomings we pose the \textit{Courtroom Challenge}: a
hybrid state machine in which every LLM claim is tested against a persistent
precedent store and—when automatic checks falter—escalated under a
tunable \textit{validation\_preference} dial (formal spec in
Section~\ref{sec:framework}).  The design rewires known parts
(generator, retrieval, human oversight) but prices each byte, cycle,
and minute of validation.  Retrieval-augmented generation \citep{lewis2020rag,
fan2024ragsurvey}—and later iterative variants—motivate the precedent
store; human-alignment work such as Constitutional AI
\citep{bai2022constitutional} motivates the escalation path.
Unlike debate frameworks \citep{chen2024reconcile,madaan2023selfrefine} or formal
provers \citep{wang2025malot}, we emphasise continuous cross-session
verification: knowledge is never final, only current.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\linewidth]{courtroom}%
      % no .pdf extension needed if LaTeX finds courtroom.pdf
  \caption{One step of the Courtroom loop.%
           \ Solid arrows: primary query flow.%
           \ Dashed arrows: helper calls and cost logging.%
           \ The cost\,bus aggregates all resource writes into a single ledger.
           \ On \textsc{DEFER} the round is skipped (no cost, no memory write).}
  \label{fig:courtroom}
\end{figure}

% ---------- quick-reference box ------------------------------------
\begin{tcolorbox}[colback=black!3,colframe=black!40,
  boxrule=0.4pt, arc=2pt, left=2pt, right=2pt,
  top=4pt, bottom=4pt, title=Notation snapshot]
\small
\begin{itemize}[leftmargin=*]
\item $G$ – \textbf{Generator}: LLM (+ optional helper calls)
that proposes a hypothesis $h_t$ for query $q_t$.
\item $V_{\theta}$ – \textbf{Validator} with
\emph{validation–preference dial} $\theta\!\in[0,1]$  
($0 =\,$speed‐first, $1 =\,$accuracy‐first).  
The dial is fixed during any evaluation run; see
Sec.~\ref{sec:framework}.
\item $\kappa_i c_{i,t}$ – \textbf{Cost-ledger term}: each resource $i$
(bytes written, CPU‐seconds, human minutes …) is logged as a
quantity $c_{i,t}$ weighted by its unit cost $\kappa_i$.
$\sum_i\kappa_i c_{i,t}$ is the \emph{friction} paid at step $t$.
\item $\tau$ – \textbf{Retention threshold} that controls whether a
new descriptor vector is written to the precedent store:  
\vspace{-2pt}
\[
\text{write\_new}=\begin{cases}
[\cos\theta < \tau], &\text{(cosine form)}\\[4pt]
[\lVert x-y\rVert_2 > \tau], &\text{(distance form)}
\end{cases}
\]
Designers may substitute any monotone similarity metric.
\end{itemize}
\vspace{2pt}\par\footnotesize (Formal definitions → Section~\ref{sec:framework})\par
\end{tcolorbox}
% --------------------------------------------------------------------

@@

\paragraph{Storage dial.}
The precedent store need not—indeed, should not—memorise every numeric
fact. A single \emph{retention knob} (\(\tau\)) decides whether a freshly
extracted descriptor vector is written back or simply reuses its nearest
neighbour. Let \(S(x,y)\) be any similarity score where “larger means
closer’’ (or equivalently a distance \(D(x,y)\) where “smaller means
closer’’). The generic rule is

@@

  \item \textbf{Memory update rule.}  
  \(U : (M, q, h, a) \to M'\) adds or merges precedent after a claim is accepted.

@@

\subsection{Information–theoretic lower bound}
\label{sec:info-bound}

\begin{lemma}[Interactive-Fano corollary]\label{lem:ErrLower}
  Fix a validator $V$ run for $T$ rounds under margin parameter $\delta>0$.
  Let
  \[
    \kappabar
      :=\mathbb{E}_{\omega}\bigl[\kappa(\omega)\bigr],
    \qquad
    \Err_{\mathrm{bar}}
      :=\mathbb{E}_{\omega}\bigl[\Err(\omega)\bigr].
  \]
  Suppose the transcript $X$ satisfies
  $I(\theta;X)\le I$
  under a uniform prior on an $M$-point label set ($M\ge 2$).
  Then
  \begin{equation}\label{eq:ErrLower}   % <-- new label
    \boxed{%
      \Err_{\mathrm{bar}}
      \;\;\ge\;\;
      \frac{\delta\,\kappabar\,T}{4}\;
      \Bigl(
        1-\dfrac{I+\log 2}{\log M}
      \Bigr)
    }.
  \end{equation}
\end{lemma}

\begin{proof}[Sketch]

\textbf{Classical identification bound.}
Choosing the 0–1 loss $L(M,X)=\mathbf 1_{\{\hat M(X)\neq M\}}$
and threshold $\Delta=\tfrac12$
in Proposition 3 of \citet{chen2024interfano} yields
\[
  P_{\mathrm{id}}
  \;\ge\;
  \tfrac12\Bigl(1-\tfrac{I+\log 2}{\log M}\Bigr).
\]

\medskip
\noindent
To obtain a bound on label error we apply the same proposition with 
threshold $\Delta=\tfrac12$ to the indicator
\[
  L^\star \;=\; \mathbf 1_{\{\Err>\Delta_{\mathrm{Err}}\}},
  \qquad
  \Delta_{\mathrm{Err}} := \tfrac12\,\delta\,\kappabar\,T.
\]

\item
  Because $\log \!\bigl(1/\rho_{1/2}\bigr)\ge\log M$ for the uniform
  prior, the proposition yields the lower bound
  \[
    \Prb\!\bigl[\Err>\Delta_{\mathrm{Err}}\bigr]
      \;\ge\;
      \frac12\Bigl(
        1-\frac{I+\log 2}{\log M}
      \Bigr).
  \]

\item
  Multiply this probability by
  $\Delta_{\mathrm{Err}}=\tfrac12\,\delta\,\kappabar\,T$
  to bound the expectation:
  \[
    \E[\Err]
      \;\ge\;
      \Delta_{\mathrm{Err}}\;
      \Prb[\Err>\Delta_{\mathrm{Err}}]
      \;=\;
      \frac{\delta\,\kappabar\,T}{4}\;
      \Bigl(
        1-\frac{I+\log 2}{\log M}
      \Bigr).
      \qedhere
  \]

@@

\paragraph{Discussion.}
Inequality~\eqref{eq:ErrLower} separates the lower bound into two
factors:

\begin{enumerate}[label=(\roman*)]
\item the \emph{size factor} $\displaystyle\frac{\delta\,\kappabar\,T}{4}$,
      which grows linearly with the horizon~$T$ and is large whenever
      the validator answers many queries ($\kappabar$ moderate) under a
      tight margin~$\delta$;
\item the \emph{information gap}
      $\displaystyle\Bigl(1-\tfrac{I+\log 2}{\log M}\Bigr)$,
      which is positive as soon as the transcript carries fewer bits
      than the label space can encode.
\end{enumerate}
Together they formalise our narrative: \emph{a low-cost, low-friction
validator cannot keep cumulative label error small indefinitely as the
stream length~$T$ grows}, unless it either (a) escalates a non-negligible
fraction of rounds (making $\kappabar$ small and thus incurring cost) or
(b) relaxes the margin~$\delta$.  In particular, when $M$ is
exponentially large and $I$ is bounded by design, $\Err_{\mathrm{bar}}$
grows linearly with~$T$.

@@

\newtheorem{conjecture}{Conjecture}
\begin{conjecture}[Validation trilemma] \label{conjecture1} Fix any open-world stream \(\mathcal{S}\) whose Shannon entropy \(\geq H\).
For every system that (i) stores \(\leq B\) bytes of precedent
and (ii) spends \(\leq \kappa\) human-minutes per 10k tokens, prequential regret
after budget B satisfies   \( R(B) \geq f(H, \kappa, B) \)   for some monotone f (non-increasing in \(\kappa\) and non-decreasing in H).
\end{conjecture}

\begin{conjecture}[No-free-lunch verifier] \label{conjecture2} Under the same entropy
assumptions, any validator \(V_\theta\) that never escalates to humans must either
(i) mis-accept adversarially crafted hallucinations with probability \(\geq \varepsilon\)
or (ii) write unbounded new precedent.
\end{conjecture}

\begin{tcolorbox}[title=Falsify-Us]
  Build an agent that—under the protocol of Section~\ref{sec:framework}—keeps
  \begin{itemize}
    \item cumulative error $\leq \varepsilon|Q|$
    \item total precedent $\leq \kappa|Q|$
    \item total human-escalation cost $\leq \beta|Q|$
  \end{itemize}
  for arbitrarily long query streams $Q$.
\end{tcolorbox}

@@

\item \emph{"But without numbers you cannot prove progress."} The claim is architectural, not empirical: \textsc{Claim A}: persistent precedent + adaptive friction is necessary for bounded-error knowledge bases in an open world; \textsc{Claim B}: current “bigger-is-better” pipelines violate \textsc{Claim A}. Both claims are falsifiable: show a practical system that achieves low hallucination without any external precedent store or escalation dial, and the Courtroom premise collapses. Until such a counter-example exists, producing another decimal improvement on GSM8K is beside the point.

\item \emph{“Isn't this just a dressed-up retrieval-augmented loop?”} Retrieval-augmented generation (RAG) treats the store as a stateless evidence buffer flushed between prompts. The Courtroom store is jurisprudential—each write becomes precedent and must pay rent (friction) forever. That difference, though conceptually small, flips the optimisation target from “accuracy at \(k = 1\)” to “prequential regret under a cost budget.” No existing RAG benchmark measures that regime.

>>>>

% ---------- math & symbols ----------
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\newtheorem*{remark}{Remark}

@@

% =========================================================
\begin{document}
\maketitle
\begin{abstract}
Large-scale generative models hallucinate, and ad hoc patches hide the true cost of verification. We pose a falsifiable ultimatum: automate validation, prove it unnecessary, or pay the explicit human-in-the-loop bill. We conjecture that for any unbounded open-world data stream with Shannon entropy rate $h>0$, a system that lacks both a persistent precedent store and a friction-priced validation loop will accumulate unbounded prequential regret. To test this “validation trilemma” we formalise a state machine, a cost ledger, and an impossibility triangle, then invite the community to break it. A 100-query toy stream illustrates how the ledger is logged; optimising the validator is left to future work. We also introduce the Courtroom Model, a wrapper that subjects every claim—textual, visual, or otherwise—to precedent search, automated semantic checks, and selective escalation to an external judge. Each output is checked against persistent vector memory, and conflicting evidence reopens prior “case law.” The challenge stands: thrive without paying the validation bill—or show why the bill is inescapable.
\end{abstract}

\footnotetext[1]{Baseline script and toy data: \url{https://github.com/johnrepsys/courtroom-model}}

\section{Introduction}
\label{sec:intro}
Large Language Models (LLMs) have demonstrated impressive capabilities in generating context-rich outputs by associating patterns from large corpora. However, this power comes with a critical limitation: LLMs often produce plausible-sounding but factually incorrect conclusions, a phenomenon known as hallucination \citep{marcus2020nextdecade,shinn2023reflexion}. Without structured verification, these errors risk being perpetuated and even amplified over time, undermining trust in LLM-generated knowledge. Addressing this issue requires a system that not only generates associations but rigorously validates and refines them.
Throughout this paper we use the term
\emph{Courtroom} metaphorically: it denotes an epistemic due-process layer that challenges and verifies LLM outputs, not a system for practicing law or adjudicating legal disputes.

To address these shortcomings we pose the \textit{Courtroom Challenge}: a
hybrid state machine in which every LLM claim is tested against a persistent
precedent store and—when automatic checks falter—escalated under a
tunable \textit{validation\_preference} dial (formal spec in
Section~\ref{sec:framework}).  The design rewires known parts
(generator, retrieval, human oversight) but prices each byte, cycle,
and minute of validation.  Retrieval-augmented generation (RAG) provides the raw materials
\citep{lewis2020rag,fan2024ragsurvey}.  
Recent advances—Self-RAG's self-verification \citep{asai2024selfrag}, 
Iter-RetGen's retrieval-generation synergy \citep{shao2023iterretgen}, 
MemoRAG's compressed long-term memory \citep{qian2024memorag}, and 
Chain-of-Verification RAG's answer-driven re-checks \citep{he2024covrag}—motivate 
our persistent precedent store but still validate only per answer.  
Human-alignment work such as Constitutional AI 
\citep{bai2022constitutional} motivates the escalation branch.  
Whereas MemoRAG retains past evidence without auditing it, our buffer makes
each accepted verdict both permanent and costly.
Unlike debate frameworks \citep{chen2024reconcile,madaan2023selfrefine} or formal
provers \citep{wang2025malot}, we emphasise continuous cross-session
verification: knowledge is never final, only current.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\linewidth]{figs/courtroom}%
      % no .pdf extension needed if LaTeX finds courtroom.pdf
  \caption{One step of the Courtroom loop.%
           \ Solid arrows: primary query flow.%
           \ Dashed arrows: helper calls and cost logging.%
           \ The cost\,bus aggregates all resource writes into a single ledger.
           \ On \textsc{DEFER} the round is skipped (no cost, no memory write).
           \ If C cannot supply an authoritative value—because the fact is genuinely unknown—the 
           \ query transitions to a terminal UNRESOLVED state; no memory write occurs.}
  \label{fig:courtroom}
\end{figure}

% ---------- quick-reference box ------------------------------------
\begin{tcolorbox}[colback=black!3,colframe=black!40,
  boxrule=0.4pt, arc=2pt, left=2pt, right=2pt,
  top=4pt, bottom=4pt, title=Notation snapshot]
\small
\begin{itemize}[leftmargin=*]
\item $G$ – \textbf{Generator}: LLM (+ optional helper calls)
that proposes a hypothesis $h_t$ for query $q_t$.
\item $V_{\theta}$ – \textbf{Validator} with
\emph{validation–preference dial} $\theta\!\in[0,1]$  
($0 =\,$speed‐first, $1 =\,$accuracy‐first).  
The dial is fixed during any evaluation run; see
Sec.~\ref{sec:framework}.
\item $\kappa_i c_{i,t}$ – \textbf{Cost-ledger term}: each resource $i$
(bytes written, CPU‐seconds, human minutes …) is logged as a
quantity $c_{i,t}$ weighted by its unit cost $\kappa_i$.
$\sum_i\kappa_i c_{i,t}$ is the friction paid at step $t$.
\item $\tau$ – \textbf{Retention threshold} that controls whether a
new descriptor vector is written to the precedent store:  
\vspace{-2pt}
\[
\text{write\_new}=\begin{cases}
[\cos\theta < \tau], &\text{(cosine form)}\\[4pt]
[\lVert x-y\rVert_2 > \tau], &\text{(distance form)}
\end{cases}
\]
Designers may substitute any monotone similarity metric.
\end{itemize}
\vspace{2pt}\par\footnotesize (Formal definitions → Section~\ref{sec:framework})\par
\end{tcolorbox}
% --------------------------------------------------------------------

@@

\paragraph{Storage dial.}
The precedent store need not—indeed, should not—memorise every numeric
fact. A single \emph{retention knob} (\(\tau\)) decides whether a freshly
extracted descriptor vector is written back or simply reuses its nearest
neighbour. Let \(S(x,y)\) be any similarity score where “larger means
closer” (or equivalently a distance \(D(x,y)\) where “smaller means
closer”). The generic rule is

@@

  \item \textbf{Memory update rule.}
  \(U : (M, q, h, a) \to M'\) records (or merges) a precedent whenever a claim is
  accepted.  Whereas vanilla RAG treats retrieved passages as a transient
  context window \citep{lewis2020rag}—and even iterative or self-reflective
  extensions still validate only the current answer
  \citep{shao2023iterretgen,asai2024selfrag}—the Courtroom store is
  persistent: it survives across sessions and every entry can later be
  re-audited for contradiction or obsolescence.

@@

\subsection{Setup}\label{sec:setup}

The definitions in \S\ref{sec:framework} merely fix notation
(\(T,\kappa,\Err,\theta,X\)); they do not restrict the validator's
internal design.  We now state Lemma~\ref{lem:ErrLower}, an
information-theoretic lower bound that applies to any interactive
validator whose transcript carries at most \(I\) bits of information
about the hidden label.  The inequality therefore functions as
motivation for the Courtroom architecture rather than a theorem
that assumes it.

\paragraph{Per-round notation.}
Fix a margin parameter $\delta\in(0,1]$.
For every run $\omega$ and round $t=1,\dots,T$ set
\begin{align*}
  y_t      &:=y_t(\theta)          &&\text{(ground-truth label)},\\
  h_t      &:=h_t(\omega)          &&\text{(validator’s automatic prediction)},\\
  \mathrm{Esc}_t&:=\mathbf 1_{\{\text{round $t$ is escalated}\}.}
\end{align*}
Define the \emph{leverage rounds}
\begin{equation}\label{eq:kappaDef}
  \mathcal L(\omega):=\Bigl\{\,t:\,
     \mathrm{Esc}_t=0\;\text{and}\;|y_t-h_t|\ge\delta\Bigr\},
  \qquad
  \kappa(\omega):=\frac{|\mathcal L(\omega)|}{T}.
\end{equation}
The cumulative label error on those rounds is
\begin{equation}\label{eq:ErrDef}
  \Err(\omega):=\sum_{t\in\mathcal L(\omega)}
                \mathbf 1_{\!\bigl[h_t\neq y_t\bigr]}.
\end{equation}
Define the global means:
\[
  \bar\kappa:=\E[\kappa],
  \qquad
  \Err_{\mathrm{bar}}:=\E[\Err].
\]

\subsection{Information-theoretic lower bound}\label{sec:info-bound}

\begin{lemma}[Interactive-Fano corollary, general prior]\label{lem:ErrLower}
Let $V$ be any validator run for $T$ rounds.
With $\kappa(\omega)$ and $\Err(\omega)$ defined in
\eqref{eq:kappaDef}-\eqref{eq:ErrDef}, write
$\Err_{\mathrm{bar}}:=\E[\Err]$.
Assume the transcript $X$ obeys the information budget
$I(\theta;X)\le I$ under an arbitrary\, prior $\mu$ on the $M$-point label set $(M\!\ge\!2)$.
Define
\[
  H_{\star}:=
  \begin{cases}
      H(\theta) & \text{(Shannon entropy)}\\[4pt]
      -\log p_{\max} & \text{(min-entropy),}
  \end{cases}
  \qquad
  p_{\max}:=\max_{m}\mu(m).
\]
Then
\begin{equation}\label{eq:ErrLower-general}
  \boxed{\,%
   \Err_{\mathrm{bar}}
   \;\ge\;
   \frac{\delta\,T}{4}\;
   \kappa_{\!*}\;
   \Bigl(1-\tfrac{I+\log 2}{H_{\star}}\Bigr),
   \quad
   \kappa_{\!*}:=
     \E\bigl[\kappa\,\bigl|\,\Err>\tfrac12\delta\kappa T\bigr]
  \,}.
\end{equation}
\end{lemma}

\begin{remark}[Entropy convention]
Throughout we write $H_{\star}$ for whichever entropy measure is appropriate
in context—Shannon entropy when a prior is specified, min-entropy when no
distributional knowledge is assumed.
Uniform priors maximise $H_{\star}$, producing the strongest (largest)
version of the bound, while skewed priors reduce $H_{\star}$ and weaken the
barrier proportionally.
\end{remark}

\begin{proof}[Proof sketch]
Step 1 (separation).
For any $m\neq m'$ and transcript $x$ at most one of
$y_t(m),y_t(m')$ can equal $h_t$ on a leverage round, hence
\[
  \Err_m(x)+\Err_{m'}(x)\;\ge\;\delta\,\kappa(\omega)\,T.
  \tag{$\ast$}
\]

\noindent
Step 2 (interactive Fano). 
Set the path-dependent loss
\[
  L^\star(m,x)
    := \mathbf{1}_{\{\Err_m(x)>\tfrac12\,\delta\,\kappa(\omega)\,T\}}
\]
and $\Delta=\tfrac12$ in Proposition 3 of
\citep{chen2024interfano}.
Inequality ($\ast$) fulfils the separation requirement
\citep[Eq.\,(5)]{chen2024interfano}.
Define
\[
  \gamma:=\frac12\Bigl(1-\tfrac{I+\log 2}{H_{\star}}\Bigr)\in(0,\tfrac12].
\]
Then
\begin{equation}\tag{$\dagger$}
  \Pr\!\bigl[\Err(\theta,X)>\tfrac12\delta\,\kappa(\omega)T\bigr]
  \;\ge\;\gamma .
\end{equation}

\medskip\noindent
Step 3 (tail $\Rightarrow$ expectation, expanded).
With
\[
  \mathcal E:=\Bigl\{\Err>\tfrac12\delta\,\kappa T\Bigr\},
  \qquad
  \kappa_{\!*}:=\E[\kappa\mid\mathcal E],
\]
we have:
\[
\begin{aligned}
  \Err_{\mathrm{bar}}
    &=\E[\Err] \\[2pt]
    &\ge
      \E\!\Bigl[
        \tfrac12\delta\,\kappa T\;
        \mathbf{1}_{\{\Err>\tfrac12\delta\,\kappa T\}}
      \Bigr] \\[2pt]
    &=\frac{\delta T}{2}\,
      \E\!\bigl[\kappa\,\mathbf{1}_{\mathcal E}\bigr] \\[2pt]
    &=\frac{\delta T}{2}\,
      \kappa_{\!*}\;
      \Pr(\mathcal E) \\[2pt]
    &\ge \frac{\delta T}{4}\,\kappa_{\!*}\, \Bigl(1-\tfrac{I+\log 2}{H_{\star}}\Bigr)\;.
\end{aligned}
\]

@@

\paragraph{Remark (relating $\kappa_{\!*}$ to $\bar\kappa$).}
Write $p_{\mathrm{bad}}:=\Pr(\mathcal E)$ with 
$\mathcal E=\{\Err>\tfrac12\delta\kappa T\}$.  
Since $0\!\le\!\kappa\!\le\!1$ one has
$\bar\kappa
   =p_{\mathrm{bad}}\kappa_{\!*}
    +(1-p_{\mathrm{bad}})\kappa^{\complement}_{\!*}
   \le p_{\mathrm{bad}}\kappa_{\!*}+1-p_{\mathrm{bad}}$.
Hence  
$\kappa_{\!*}\ge\bar\kappa+p_{\mathrm{bad}}-1
            \ge\bar\kappa+\gamma-1$,
where $\gamma$ is the Fano factor.  Therefore
$\kappa_{\!*}\ge\max\{0,\bar\kappa+\gamma-1\}$ as used in the following
discussion.

\paragraph{Discussion.}
Write the bound as  
\(\Err_{\mathrm{bar}}\!\ge(\tfrac12\delta T)\,\max\{0,\bar\kappa+\gamma-1\}\).
Because \(\gamma>0\) whenever the transcript leaks
fewer than \((H_{\star}-\log 2)\) bits, any validator that
handles a non-trivial share of rounds automatically
\((\bar\kappa>1-\gamma)\)
faces a linear-in-$T$ unavoidable error floor.
Escalating a larger fraction of rounds (making $\bar\kappa$ smaller) or
relaxing the margin $\delta$ are the only ways to suppress this growth,
precisely matching the “friction buys accuracy” narrative.

@@

\newtheorem{conjecture}{Conjecture}
\begin{conjecture}[Validation trilemma]\label{conjecture1}
Let $\mathcal{S}=\langle s_1,\dots,s_T\rangle$ be an open-world data stream
whose source has Shannon entropy rate at least $h$ bits per symbol
(equivalently, $H(\mathcal S_{1:T})\ge hT$).
For any system that\\[0.1em]
\hspace*{1.5em}(i) stores at most $B$ bytes of persistent precedent and\\[0em]
\hspace*{1.5em}(ii) budgets at most $\kappa$ bits of external validation per $N$ symbols,\\
the prequential regret after it consumes $B$ bytes satisfies
\[
   R(B)\;\;\ge\;\;f\bigl(h,\kappa,B\bigr),
\]
for some monotone $f$ (non-increasing in $\kappa$, non-decreasing in $h$).
\end{conjecture}

\begin{conjecture}[No-free-lunch verifier]\label{conjecture2}
Under the same entropy assumption,
any fully automatic validator $V_\theta$ that never escalates to humans
must either
\begin{enumerate}[label=(\roman*)]
\item mis-accept adversarially crafted claims with probability at least~$\varepsilon$, or
\item write an unbounded amount of new precedent as $T\to\infty$.
\end{enumerate}
\end{conjecture}

\begin{tcolorbox}[title=Falsify-Us]
  Build an agent that—under the protocol of Section~\ref{sec:framework}—keeps
  \begin{itemize}
    \item cumulative error $\leq \varepsilon|Q|$
    \item total precedent $\leq \kappa|Q|$
    \item total human-escalation cost $\leq \beta|Q|$
  \end{itemize}
  for arbitrarily long open-world data streams $S$.
\end{tcolorbox}

@@

\item \emph{"But without numbers you cannot prove progress."} The claim is architectural, not empirical: \textsc{Claim A}: persistent precedent + adaptive friction is necessary for bounded-error knowledge bases in an open world; \textsc{Claim B}: current “bigger-is-better” pipelines violate \textsc{Claim A}. Both claims are falsifiable: show a practical system that achieves low hallucination without any external precedent store or escalation dial, and the Courtroom premise collapses. Until such a counter-example exists, producing another decimal improvement on the latest benchmark leaderboard is beside the point.

\item \emph{“Isn't this just a dressed-up retrieval-augmented loop?”} Retrieval-augmented generation (RAG) treats the store as a context-window scoped evidence buffer flushed between prompts. The Courtroom store is jurisprudential—each write becomes precedent and must pay rent (friction) forever. That difference, though conceptually small, flips the optimisation target from “accuracy at \(k = 1\)” to “prequential regret under a cost budget.” No existing RAG benchmark measures that regime.
